---
title: "Class_1_R_Rev"
author: "Rubagumya"
date: "`r Sys.Date()`"
output: html_document
---

## Data Structure(Vectors,Factors,Lists,DataFrame, matrix and arrays)
### Vectors: The foundational R data structure
#### Numeric
```{r}
x<- c(0.5,0.6)
age<-c(20,35,32,29)
summary(age)
```
```{r}

```


## Logical
```{r}
x <- c(TRUE,FALSE)
x
```
```{r}
x <- c(T,F,T)
x
```
## Character
```{r}
x<- c('a','b')
x
class <- c('M','F','M')
class
```
##Integer
```{r}
x<- c(9:29)
x
```
##Comple numbers
```{r}
x<- c(1+2i,2+1i)
x
class(x)
```
## Matrix and operation within matrices and dim
### Using Matrix(data, nrow = rows, ncol = cols)
```{r}
x<- matrix(1:6, nrow = 3, ncol = 2)
x
class(x)

y<- c(1:12)
z<- matrix(y,nrow = 3,ncol = 3)
z
class(z)
```

### dim(data) <-(nrow,ncol)
```{r}
dim(y) <- c(3,4)
y
```
### Matrix operations
```{r}
x<- matrix(1:6, nrow = 2, ncol = 3)     #Initializing matrix x
y<- matrix(1:6, nrow = 3, ncol = 2)    #Initializing matrix y
#z

#Matrix multiplication xXy
print('Multiplication of x*y')
print(x%*%y)
print('Multiplication of x*Z')
print(x%*%z)
print('Multiplication of y*z')
#print(y%*%z)
```
### Lists
lists can be created using List() func
```{r}
x <- list(1,'a',TRUE,1+0i)
x
```
###We cam crate an empty list with vector() function
```{r}
x<- vector('list',length = 5)
x
```
## Factors

```{r}
x<- factor(c('yes','yes','no','no'))
x
```
## Table
```{r}
table(x)
```
##The underlying representation of factor
```{r}
unclass(x)
```
```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"))
x ## Levels are put in alphabetical order

```
## Missing values 
```{r}
x<- c(1,2,NA,10,3)
x
## Return logical indicating which elements are NA
is.na(x)
```
```{r}
x <- c(1, 2, NaN, NA, 4)   # a vector with both na and Nan
is.na(x)
is.nan(x)
```
## Coercion

```{r}
x <- 1:9
class(x)
as.numeric(x)
as.character(x)
as.logical(x)
```
## Creating data
### using Data frame
```{r}
studentID<-c(1,2,3,4,5)
math_score <- c(12,17,10,9,NA)
gender<-c('M','F','M','M','F')
it_score<-c(13,18,11,10,19)
scoredata<-data.frame(studentID,gender,math_score,it_score)
scoredata
```
## Creating data from keyboard
```{r}
#dt_class<- data.frame(height = numeric(0),weight = numeric(0), bmi = numeric(0))
#dt_class<-edit(dt_class)
```

## Data Importation
### Data from excel
```{r}
df_cars_sales<- read.csv('cars_data.csv')
head(df_cars_sales)
```

### Head data
```{r}
head(df_cars_sales)
```

### Tail data
```{r}
tail(df_cars_sales)
```
### Slicing the df

```{r}
df_cars_sales[10:20,]
```

### Summary of the dataset
```{r}
summary(df_cars_sales$Mileage)
```
```{r}
length(df_cars_sales$Mileage)
```

```{r}
library(haven)
```


### importing dta file
```{r}
dt_dta_file <- read_dta('stat_packages.dta')
head(dt_dta_file)
View(dt_dta_file)
```

### Importing sas file
```{r}
df_sas_file<- read_sas('stat_packages.sas7bdat')
head(df_sas_file)
```
### Importing sav file  

```{r}
df_sav_file<- read_sav('stat_packages.sav')
df_sav_file
```
### Import Xml file 

```{r}
library(xml2)
df_xml_file<- read_xml('students.xml')
df_xml_file
```
### Import sql
```{r}
df_sql_file<- readLines('students.sql')
df_sql_file
```


